# (PART\*) Two-Way ANOVA {-}

# Data Entry^[The contents of the succeeding sections are obtained from this [link](http://www.sthda.com/english/wiki/two-way-anova-test-in-r#compute-two-way-anova-test-in-r-for-unbalanced-designs)]

Load data to R using `read_csv()` function of the `readr` package of `tidyverse` and save it with a variable name `twoway_data`.


```{r message=FALSE, warning=FALSE}
# Load and save
twoway_data <- read_csv(file = "data/Tubo-USEP_Two-Way Cleaned Data for R.csv")
# Preview
twoway_data
```


# Data Manipulation

```{r}
twoway_data%>% 
  gather(key = Plot, value = Yield, -c(Fertilizer,Manure)) %>% 
  mutate(across(Fertilizer:Plot, ~ as.factor(.x))) -> clean_twoway_data
```
```{r}
# Structure preview
str(clean_twoway_data)
```

```{r}
# Sample preview
clean_twoway_data %>% 
  sample_n(10)
```

# Data Visualization

```{r}
ggboxplot(clean_twoway_data,
          x = "Fertilizer",
          y = "Yield",
          fill = "Manure")
```

# Hypothesis Testing

## The Two-Way ANOVA Table with Main Effects Only

```{r}
two_aov <- clean_twoway_data %>% 
  aov(formula = Yield ~ Fertilizer + Manure, data = .)
  
two_aov %>% 
  summary
```


Similar to the One-Way ANOVA table, 
  
  - `Df` -- degress of freedom
  - `Sum Sq` -- sum of squares
  - `Mean Sq` -- mean sum of squares
  - `F value` -- value of \(F\) statistic
  - `Pr(>F)` -- \(p\)-value

Thus, from the table

  \begin{align}
    SSR &= 17.67 & MSR &= 17.672 & F_C = 6.332\\
    SSC &= 19.21 & MSC &= 19.208 & F_R = 6.883\\
    SSE &= 47.44 & MSE &= 2.791  & \\
  \end{align}

Similar to when you look up at an F-table, the p-values can be computed using the following R code.

```{r}
pf(q = 6.332, df1 = 1, df2 = 17, lower.tail = F)
pf(q = 6.883, df1 = 1, df2 = 17, lower.tail = F)
```

## The Two-Way ANOVA Table with Interactions

```{r}
two_aov2 <- clean_twoway_data %>% 
  aov(formula = Yield ~ Fertilizer*Manure, data = .)
  
two_aov2 %>% 
  summary
```

The interaction is not significant so we proceed on using the additive model (i.e., main-effects only).

# Checking Assumptions


## Checking Normality Assumptions

_Shapiro-Wilk Test_

The Shapiro-Wilk test tests the null hypothesis that the samples come from a normal distribution against the alternative hypothesis that the samples do not come from a normal distribution. 

```{r}
shapiro.test(residuals(two_aov))
```
_QQ Plots_

```{r}
plot(two_aov, 2)
```

## Checking Homogeneity of Variance Assumption

```{r}
par(mfrow=c(2,2))
plot(two_aov)
par(mfrow=c(1,1))
```

<!-- # Post-hoc -->
<!-- We no longer need a posthoc analysis since we only have two levels in each of the significant factors. However, for the sake of illustration, we could be simply run the following code -->

<!-- ```{r} -->
<!-- TukeyHSD(two_aov) -->
<!-- ``` -->

# Mean Line or Interaction Plots

```{r}
ggline(clean_twoway_data,
       x = "Fertilizer",
       y = "Yield",
       color = "Manure",
       add = c("mean_se"))
```

The above figure supports the result before that there is no significant interaction between the two factors.